On exponential sums arising from the classical doubling method

### 2025-03-20 paper.tex
Macros

% Finite field macros
\newcommand{\FieldNorm}[2]{\mathrm{N}_{#1:#2}}
\newcommand{\aFieldNorm}{\mathrm{N}}
\newcommand{\finiteField}{\mathbb{F}}
\newcommand{\quadraticExtension}{\mathbb{E}}
\newcommand{\finiteFieldExtension}[1]{\finiteField_{#1}}
\newcommand{\quadraticFieldExtension}[1]{\quadraticExtension_{#1}}
\newcommand{\NormOneGroup}[1]{\finiteFieldExtension{#1}^{\aFieldNorm = 1}}
\newcommand{\algebraicClosure}[1]{\overline{#1}}
\newcommand{\Galois}{\operatorname{Gal}}
\newcommand{\Frobenius}{\operatorname{Fr}}
\newcommand{\restrictionOfScalars}[3]{\operatorname{Res}_{#1 \slash #2}{#3}}
\newcommand{\multiplcativeScheme}{\mathbb{G}_m}
\newcommand{\affineLine}{\mathbb{A}^1}
\newcommand{\squareMatrix}{\operatorname{Mat}}
\newcommand{\Mat}[2]{\operatorname{Mat}_{#1 \times #2}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSum}[2]{\mathcal{J}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}

@@

\subsubsection{Geometric Conjugacy and Lusztig Series}
Let $\algebraicGroup{G}$ be a reductive group over $\finiteField$ and fix a torus $\algebraicGroup{T}$ defined over $\finiteField$. The \emph{dual group over $\finiteField$} of $\algebraicGroup{G}$ with respect to the torus $\algebraicGroup{T}$ corresponds to the group $\algebraicGroup{G}^*$ with a dual torus $\algebraicGroup{T}^*$ and root datum
\[
    (X^*(\algebraicGroup{T}^*),\Phi_{\algebraicGroup{T}^*},X_*(\algebraicGroup{T}^*),\Phi_{\algebraicGroup{T}^*}^\vee) \cong (X_*(\algebraicGroup{T}),\Phi_{\algebraicGroup{T}}^\vee,X^*(\algebraicGroup{T}),\Phi_{\algebraicGroup{T}})
\]
such that the isomorphism respects the Frobenius action on $X^*$ and $X_*$.

Fix an isomorphism $\multiplicativegroup{\algebraicClosure{\finiteField}} \cong (\mathbb{Q}/\mathbb{Z})_{p'}$ and an embedding $\exp:(\mathbb{Q}/\mathbb{Z})_{p'} \to \overline{\mathbb{Q}_\ell}^\times$. 
We get an isomorphism
\begin{equation}
\label{eq:DL}
\xymatrix{0 \ar[r] & X_*(\algebraicGroup{T})_{(\mathbb{Q}/\mathbb{Z})_{p'}} \ar[r]^{\Frobenius-\idmap} & X_*(\algebraicGroup{T})_{(\mathbb{Q}/\mathbb{Z})_{p'}} \ar[r] & \algebraicGroup{T}(\algebraicClosure{\finiteField}) \ar[r] & 0}
\end{equation}
with the last map being $y \mapsto \aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(y(\zeta_n))$
for $\zeta_n$ being a $q^n-1$ root of unity corresponding to $1/(q^n-1) \in (\mathbb{Q}/\mathbb{Z})_{p'}$ and sufficiently large $n$ (large enough so that $\algebraicGroup{T}\left(\finiteFieldExtension{n}\right)$ is split). Here, $$\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(t) = t\cdot \Frobenius(t)\cdot \Frobenius^2(t)\cdot  \hdots \cdot \Frobenius^{n-1}(t)$$ is the torus norm map.

Hence for a torus character pair $(T,\theta)$ for $G$, we may conjugate $\algebraicGroup{T}$ to a fixed torus $\algebraicGroup{T}'$ and associate to $\theta$ a twisted Frobenius (twisted by an element of $W$ in the twisted Weyl group conjugacy class corresponding to $T$) stable element $x_{T,\theta}$ of $X^*(\algebraicGroup{T}')_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ via the equality
\[
    \exp(\langle x_{T,\theta},y\rangle) = \theta(\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(y(\zeta_n)))
\]
for large enough $n$. This $x_{T,\theta}$ is well defined up to the Weyl group, and in fact, by \cite[Corollary 13.9]{DigneMichel2020} there is a correspondence between $(T,\theta)$ pairs and $\Frobenius$-stable Weyl group orbits in $X^*(\algebraicGroup{T}')_{(\mathbb{Q}/\mathbb{Z})_{p'}}$.  By composing with \eqref{eq:DL}, we associate to a torus character pair a Frobenius stable semisimple conjugacy class in $G^*$.

    Denote the Frobenius semisimple conjugacy class associated to a torus character pair $(T,\theta)$ by $(s_{T,\theta})$. Similarly denote the Frobenius stable Weyl group element of $X^*(\algebraicGroup{T})_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ corresponding to $(T,\theta)$ by $(x_{T,\theta}).$

>>>>

% Finite field macros
\newcommand{\FieldNorm}[2]{\mathrm{N}_{#1:#2}}
\newcommand{\aFieldNorm}{\mathrm{N}}
\newcommand{\finiteField}{\mathbb{F}}
\newcommand{\quadraticExtension}{\mathbb{E}}
\newcommand{\finiteFieldExtension}[1]{\finiteField_{#1}}
\newcommand{\quadraticFieldExtension}[1]{\quadraticExtension_{#1}}
\newcommand{\NormOneGroup}[1]{\finiteFieldExtension{#1}^{\aFieldNorm = 1}}
\newcommand{\algebraicClosure}[1]{\overline{#1}}
\newcommand{\Galois}{\operatorname{Gal}}
\newcommand{\Frobenius}{\operatorname{Fr}}
\newcommand{\restrictionOfScalars}[3]{\operatorname{Res}_{#1 \slash #2}{#3}}
\newcommand{\multiplcativeScheme}{\mathbb{G}_m}
\newcommand{\affineLine}{\mathbb{A}^1}
\newcommand{\squareMatrix}{\operatorname{Mat}}
\newcommand{\Mat}[2]{\operatorname{Mat}_{#1 \times #2}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSum}[2]{\mathcal{J}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblJacobiSumScalar}[2]{\mathrm{J}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualJacobiSumScalar}[2]{\mathrm{j}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\dblLanglandsGammaFactorSpace}[4]{\gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}
\newcommand{\LusztigSeries}[2]{\mathcal{E}\left(#1, (#2)\right)}
\newcommand{\characteristicPolynomial}{\operatorname{CharPoly}}
\newcommand{\DualFrobeniusFixedPoints}[2][\Frobenius^{\ast}]{\algebraicGroup{#2}^{\ast #1}}
\newcommand{\FrobeniusFixedPoints}[2][\Frobenius]{\algebraicGroup{#2}^{#1}}
\newcommand{\CharacterLattice}[1]{X^{\ast}\left(#1\right)}
\newcommand{\CocharacterLattice}[1]{X_{\ast}\left(#1\right)}

@@

\subsubsection{Geometric Conjugacy and Lusztig Series}
Let $\algebraicGroup{G}$ be a reductive group over $\finiteField$ and fix a torus $\algebraicGroup{T}$ defined over $\finiteField$. The \emph{dual group over $\finiteField$} of $\algebraicGroup{G}$ with respect to the torus $\algebraicGroup{T}$ corresponds to the group $\algebraicGroup{G}^*$ with a dual torus $\algebraicGroup{T}^*$ and root datum
\[
    (\CharacterLattice{\algebraicGroup{T}^*},\Phi_{\algebraicGroup{T}^*},\CocharacterLattice{\algebraicGroup{T}^*},\Phi_{\algebraicGroup{T}^*}^\vee) \cong (\CocharacterLattice{\algebraicGroup{T}},\Phi_{\algebraicGroup{T}}^\vee,\CharacterLattice{\algebraicGroup{T}},\Phi_{\algebraicGroup{T}})
\]
such that the isomorphism respects the Frobenius actions on $\CharacterLattice{\cdot}$ and $\CocharacterLattice{\cdot}$.

Fix an isomorphism $\multiplicativegroup{\algebraicClosure{\finiteField}} \cong (\mathbb{Q}/\mathbb{Z})_{p'}$ and an embedding $\exp:(\mathbb{Q}/\mathbb{Z})_{p'} \to \overline{\mathbb{Q}_\ell}^\times$. 
By \cite[Proposition 13.7 (ii)]{DigneMichel2020}, we have an exact sequence
\begin{equation}
\label{eq:DL}
\xymatrix{0 \ar[r] & \CocharacterLattice{\algebraicGroup{T}}_{(\mathbb{Q}/\mathbb{Z})_{p'}} \ar[r]^{\Frobenius-\idmap} & \CocharacterLattice{\algebraicGroup{T}}_{(\mathbb{Q}/\mathbb{Z})_{p'}} \ar[r] & \algebraicGroup{T}(\algebraicClosure{\finiteField}) \ar[r] & 0}
\end{equation}
with the last map being $y \mapsto \aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(y(\zeta_n))$
for $\zeta_n$ being a $q^n-1$ root of unity corresponding to $1/(q^n-1) \in (\mathbb{Q}/\mathbb{Z})_{p'}$ and sufficiently large $n$ (large enough so that $\algebraicGroup{T}\left(\finiteFieldExtension{n}\right)$ is split). Here, $$\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(t) = t\cdot \Frobenius(t)\cdot \Frobenius^2(t)\cdot  \hdots \cdot \Frobenius^{n-1}(t)$$ is the torus norm map.

Hence for a torus character pair $(T,\theta)$ for $G$, we may conjugate $\algebraicGroup{T}$ to a fixed torus $\algebraicGroup{T}'$ and associate to $\theta$ a twisted Frobenius (twisted by an element of $W$ in the twisted Weyl group conjugacy class corresponding to $T$) stable element $x_{T,\theta}$ of $\CharacterLattice{\algebraicGroup{T}'}_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ via the equality
\[
    \exp(\langle x_{T,\theta},y\rangle) = \theta(\aFieldNorm_{\algebraicGroup{T}^{\Frobenius^n}:\algebraicGroup{T}^{\Frobenius}}(y(\zeta_n)))
\]
for large enough $n$. This $x_{T,\theta}$ is well defined up to the Weyl group, and in fact, by \cite[Corollary 13.9]{DigneMichel2020} there is a correspondence between $(T,\theta)$ pairs and $\Frobenius$-stable Weyl group orbits in $\CharacterLattice{\algebraicGroup{T}'}_{(\mathbb{Q}/\mathbb{Z})_{p'}}$.  By composing with \eqref{eq:DL}, we associate to a torus character pair a Frobenius stable semisimple conjugacy class in $G^*$.

    Denote the Frobenius semisimple conjugacy class associated to a torus character pair $(T,\theta)$ by $(s_{T,\theta})$. Similarly, denote the Frobenius stable Weyl group element of $\CharacterLattice{\algebraicGroup{T}}_{(\mathbb{Q}/\mathbb{Z})_{p'}}$ corresponding to $(T,\theta)$ by $(x_{T,\theta}).$
